\section{Status}

\paragraph*{Status of GSL-Library}
The gsl-library is since version 1.0 stable and for general use.
More information about it at \url{http://www.gnu.org/software/gsl/}.

\paragraph*{Status of this interface}
Nearly all modules are wrapped. A lot of tests are
covering various functionality. Please report to the mailing list
\url{pygsl-discuss@lists.sourceforge.net} if you find a bug.

The hankel modules have been
wrapped. Please write to the mailing list
\url{pygsl-discuss@lists.sourceforge.net} 
if you require one of the modules
and are willing to help with a simple example. 
If any other function is missing or some other module (e.g. ntuple) or
function, do not hesitate to write to the list.

\paragraph*{Retriving the Interface}
You can download it here: \url{http://sourceforge.net/projects/pygsl}

\section{Requirements}

To build the interface, you will need
\begin{itemize}
\item \ulink{gsl-1.x}{http://sources.redhat.com/gsl},
\item \ulink{python2.2}{http://www.python.org} or better,
\item \ulink{NumPy}{http://numpy.sf.net}, and
\item a c compiler (like \ulink{gcc}{http://gcc.gnu.org}).
\end{itemize}

Supported Platforms are:
\begin{itemize}
\item Linux (Redhat/Debian/SuSE) with python2.* and gsl-1.*
\item Win32
\end{itemize}
It was tested and is tested on an irregular basis on the following platforms
\begin{itemize}
\item SUN
\item Cygwin
\item MacOS X
\end{itemize}
but is supposed to build on any POSIX platforms.

\section{Installing the pygsl interface}

\program{gsl-config} must be on your path:\nopagebreak
\begin{verbatim}
# unpack the source distribution
gzip -d -c pygsl-x.y.z.tar.gz|tar xvf-
cd pygsl-x.y.z
# do this with your prefered python version
# to set the gsl location explicitly use setup.py --gsl-prefix=/path/to/gsl
python setup.py build
# change to an user id, that is allowed to do installation
python setup.py install
\end{verbatim}
Ready....

{\bf Do not test the interface in the distribution root or in the directories
 \file{src} or \file{pygsl}.}

\subsection{Building on win32}

Windows by default does not allow to run a posix shell. Here a different path
is required. First change into the directory \file{gsl_dist}. Copy the file 
\file{gsl_site_example.py}
and edit it to reflect your installation of GSL and SWIG if you want to run it
yourself. The pygsl windows binaries distributed over 
\url{http://sourceforge.net/projects/pygsl/} are built using the mingw32 
compiler. 

\paragraph*{Uninstall GSL interface}
\code{rm -r }"python install path"\code{/lib/python}"version"\code{/site-packages/pygsl}

\paragraph*{Testing}
the directory \file{tests} contains several testsuites, based on python
\module{unittest}.
The script \file{run_test.py} in this directory will run one after the other.

\paragraph*{Support}
Please send mails to our mailinglist at
\email{pygsl-discuss@lists.sourceforge.net}.

\paragraph*{Developement}
You can browse our cvs tree at
\url{http://cvs.sourceforge.net/cgi-bin/viewcvs.cgi/pygsl/pygsl/}.
\\
Type this to check out the actual version:
\begin{verbatim}
cvs -d:pserver:anonymous@cvs.pygsl.sourceforge.net:/cvsroot/pygsl login
#Hit return for no password.
cvs -z3 -d:pserver:anonymous@cvs.pygsl.sourceforge.net:/cvsroot/pygsl co pygsl
\end{verbatim}
The script \program{tools/extract_tool.py} generates most of the special 
function code.

%\input{install_advanced.tex}
\paragraph*{ToDo}
Implement other parts:


\paragraph*{History}
\begin{itemize}
\item a gsl-interface for python was needed for a project at
\ulink{Center for Applied Informatics Cologne}{http://www.zaik.uni-koeln.de/AFS}.
\item \file{gsl-0.0.3} was released at May 23, 2001
\item \file{gsl-0.0.4} was released at January 8, 2002
\item \file{gsl-0.0.5} is growing since January, 2002
\item \file{gsl-0.2.0} was released at 
\item \file{gsl-0.3.0} was released at 
\item \file{gsl-0.3.1} was released at 
\item \file{gsl-0.3.2} was released at 
\item \file{gsl-0.9.4} was released at 25. October 2008
\end{itemize}

\paragraph*{Thanks}
Jochen K\"upper (\email{jochen@jochen-kuepper.de}) for 
\module{pygsl.statistics} part\\
Fabian Jakobs for \module{pygsl.blas}, \module{pygsl.eigen}
\module{pygsl.linalg}, \module{pygsl.permutation}\\ 
Leonardo Milano for rpm build\\
Eric Gurrola and  Peter Stoltz for testing and supporting the port of pygsl to
the MAC\\
Sebastien Maret for supporting the Fink \url{http://fink.sourceforge.net}
port of pygsl.


\paragraph*{Maintainers}
Achim G\"adke (\email{AchimGaedke@users.sourceforge.net}),\\
Pierre Schnizer (\email{schnizer@users.sourceforge.net})
