\chapter[\protect\module{pygsl.testing} ---  Modules in Testing]
{\protect\module{pygsl.testing} \\ Modules in Testing}
\label{cha:statistics-module}

\declaremodule{standard}{pygsl.testing}

\moduleauthor{Pierre Schnizer}{schnizer@users.sourceforge.net}
Modules in this package are often reimplementations of an original package
with significant change to the original. The current rng implementation, for
example, started its life here.  The sf module implemented here will supersede
the sf package in one of the next releases. Concerning the other modules the 
usage is encouraged for tests
to see if they work, but use them with caution in your production code!

\section[\protect\module{pygsl.testing.sf} --- Special UFuncs]
{\protect\module{pygsl.testing.sf} \\ Special Functions as UFuncs}

\declaremodule{standard}{pygsl.testing.sf}
\moduleauthor{Pierre Schnizer}{schnizer@users.sourceforge.net}

This chapter provides mainly \numpy{} UFuncs over the special functions. This means
that all input variable can be arrays, and the UFunc will evaluate the gsl
function for all its inputs. It is meant to replace the sf module later;
please use it and find out if it is useful for you. 
Only the python specific part is described here. For a general description of
the function please see the GSL Reference document.  

\section{UFuncs}
\label{sec:ufuncs}
These UFuncs allow to evaluate an array of doubles or an array of floats
typically. Note that the return values of the functions ending with an _e are
not turned to exceptions. Instead the error codes are returned as an separate
array. So if you are interested in them you have to do check the error code
yourself. 

\input{providedufuncs.tex}

\section{Ordinary Functions}

The following array functions have been wrapped. These are supposingly faster
than the equivalent functions from above.
\begin{funcdesc}{bessel_In_array}{...}\index{bessel_In_array}
\end{funcdesc}
\begin{funcdesc}{bessel_Jn_array}{...}\index{bessel_Jn_array}
\end{funcdesc}
\begin{funcdesc}{bessel_Kn_array}{...}\index{bessel_Kn_array}
\end{funcdesc}
\begin{funcdesc}{bessel_Kn_scaled_array}{...}\index{bessel_Kn_scaled_array}
\end{funcdesc}
\begin{funcdesc}{bessel_Yn_array}{...}\index{bessel_Yn_array}
\end{funcdesc}
\begin{funcdesc}{bessel_il_scaled_array}{...}\index{bessel_il_scaled_array}
\end{funcdesc}
\begin{funcdesc}{bessel_jl_array}{...}\index{bessel_jl_array}
\end{funcdesc}
\begin{funcdesc}{bessel_jl_steed_array}{...}\index{bessel_jl_steed_array}
\end{funcdesc}
\begin{funcdesc}{bessel_kl_scaled_array}{...}\index{bessel_kl_scaled_array}
\end{funcdesc}
\begin{funcdesc}{bessel_yl_array}{...}\index{bessel_yl_array}
\end{funcdesc}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "ref"
%%% End: 
